---
title: "Water ORGs data cleaning"
toc: true
format:
  html: 
    code-fold: true
    theme: simplex
---

# 0) Load packages and data

**Packages**

```{r}
pacman::p_load(tidyverse, here, sf, readxl, janitor, rio, DataExplorer, fuzzyjoin, tidygeocoder)
```

# 1) Create a dataset with water organizations list by country

## 1.1) Load data

```{r, warning=FALSE, message=FALSE, results='hide'}
wwf_capacity <- read_xlsx(here("input", "water_orgs", "WWF_capacity.xlsx"))

other_ngos <- read_xlsx(here("input", "water_orgs", "Collective Action Actor Mapping - Dec 2020.xlsx"))

country <- st_read(here("input", "geographies", "world_adm0_esri_edited_diss.shp"))
```

## 1.2) Clean data

### 1.2.1) WWF Capacity

```{r}
wwf_capacity <- wwf_capacity %>% 
  distinct() %>% 
  mutate(Country = case_when(
    Country == "Republic of Korea" ~ "South Korea",
    Country == "UAE" ~ "United Arab Emirates",
    Country == "DRC" ~ "Congo DRC",
    Country == "Guianas" ~ "Guayana",
    Country == "Russia" ~ "Russian Federation",
    TRUE ~ Country
  ))
```

### 1.2.2) Other NGOs

```{r}
other_ngos <- other_ngos %>% 
  row_to_names(1) %>% 
  remove_empty(c("rows", "cols")) %>% 
  rename(capacity = `Capacity (0 = no water staff; 1 = 1-5 staff; 2 = 6-50 staff; 3 = >50 staff)`,
         `South Korea` = `Korea, Republic of`,
         Kyrgyzstan = `Kyrgyz Republic`,
          Laos =  `Lao People's Democratic Republic`,
         Micronesia = `Micronesia, Federated States of`,
         Vietnam = `Viet Nam`)
```

## 1.3) Join datasets together

```{r}
other_ngos_capacity <- other_ngos %>% 
  dplyr::select(Organisations, 
                capacity, 
                Afghanistan:Vietnam) %>% 
  pivot_longer(cols = Afghanistan:Vietnam,
               names_to = "country",
               values_to = "ngo_presence") %>% 
  set_missing("0") %>% 
  mutate(capacity = as.numeric(capacity) * (as.numeric(ngo_presence)))
```

```{r}
df <- other_ngos %>% 
  dplyr::select(Organisations, 
                Afghanistan:Vietnam) %>%
  mutate(Organisations = case_when(
    Organisations == "2030 Water Resources Group (2030WRG)" ~ "2030WRG",                           
    Organisations == "Alliance for Global Water Adaptation" ~ "AGWA",                            
    Organisations == "Alliance for Water Stewardship (AWS)" ~ "AWS",                            
    Organisations == "Better Cotton Initiative (BCI)" ~ "BCI",                                  
    Organisations == "Beverage Industry Environment Roundtable (BIER)" ~ "BIER",                 
    Organisations == "CDP" ~ "CDP",                                                             
    Organisations == "CEO Water Mandate" ~ "CEOWM",                                               
    Organisations == "Ceres" ~ "Ceres",                                                           
    Organisations == "China Water Risk (CWR)" ~ "CWR",                                          
    Organisations == "Financial Centre for South-South Cooperation" ~ "FCSSC",                    
    Organisations == "GIZ (NatuRES)" ~ "GIZ",                                                   
    Organisations == "Global Environment Agency (GEF)" ~ "GEF",                                 
    Organisations == "Global Environment Management Initaitive (GEMI)" ~ "GEMI",                 
    Organisations == "Global Water Challenge (GWC)" ~ "GWC",                                    
    Organisations == "Global Water Partnership (GWP)" ~ "GWP",                                  
    Organisations == "Good Stuff International" ~ "GSI",                                        
    Organisations == "Greenpeace" ~ "Greenpeace",                                                      
    Organisations == "International Council on Mining & Metals (ICMM)" ~ "ICMM",                 
    Organisations == "International Water Association (IWA)" ~ "IWA",                           
    Organisations == "International Water Management Institute (IWMI)" ~ "IWMI",                 
    Organisations == "International Water Resources Association (IWRA)" ~ "IWRA",                
    Organisations == "IPE" ~ "IPE",                                                             
    Organisations == "IUCN" ~ "IUCN",                                                            
    Organisations == "Lien AID" ~ "LienAID",                                                        
    Organisations == "Natural Resources Defense Council (NRDC)" ~ "NRDC",                        
    Organisations == "OECD" ~ "OECD",                                                            
    Organisations == "Pacific Institute" ~ "PI",                                               
    Organisations == "RVO (Water As Leverage fro Resilient Asian Cities program - WaL)" ~ "RVO",
    Organisations == "Sustainable Apparel Coalition (SAC)" ~ "SAC",                             
    Organisations == "Sustainable Rice Platform" ~ "SRP",                                       
    Organisations == "Swedish Textile Water Initiative (STWI)" ~ "STWI",                         
    Organisations == "The Nature Conservancy (TNC)" ~ "TNC",                                    
    Organisations == "Water for People" ~ "WforPeople",                                                
    Organisations == "Water Foundry" ~ "WFoundry",                                                   
    Organisations == "Water Witness International (WWI)" ~ "WWI",                               
    Organisations == "Water.org" ~ "W.org",                                                       
    Organisations == "WaterAid" ~ "WaterAid",                                                        
    Organisations == "World Business Council for Sustainable Development (WBCSD)" ~ "WBCSD",      
    Organisations == "World Resource Institute (WRI)" ~ "WRI",                                  
    Organisations == "WWF" ~ "WWF" 
  ))
```

```{r}
col_nam <- (df %>% names)[-1]

country_org <- col_nam %>% 
  map_df(function(country){
    
    df2 <- df %>% 
      dplyr::select(1, country) %>% 
      drop_na(country)
    
    tibble(country = country, organizations = df2$Organisations %>% paste(collapse=", "))
  })
```

```{r}
ngo_country <- country %>% 
  stringdist_left_join(
    other_ngos_capacity,
    by = c("NAME" = "country"),
    method = "lv",
    max_dist = 1) %>% 
  st_drop_geometry() %>% 
  dplyr::select(-c(ISO_2DIGIT, ISO_3DIGIT, country, Area_km2)) %>% 
  rename(country = NAME) %>% 
  mutate(ngo_presence = ifelse(
    is.na(ngo_presence),
    0,
    ngo_presence),
    capacity = as.numeric(capacity),
    ngo_presence = as.numeric(ngo_presence)) %>% 
  group_by(country) %>% 
  summarise(capacity_sum = sum(capacity, na.rm = T),
            ngo_presence = sum(ngo_presence, na.rm = T),
            n_org = sum(ngo_presence, na.rm = T)
  ) %>% 
  mutate(ngo_presence = ifelse(
    ngo_presence > 1,
    "yes",
    "no"
  )) %>% 
  left_join(country_org) %>% 
  arrange(country) %>% 
  stringdist_left_join(
    wwf_capacity,
    by = c("country" = "Country"),
    method = "lv",
    max_dist = 1) %>% 
  dplyr::select(-Country) %>% 
  dplyr::select(country, ngo_presence, n_org, capacity_sum, organizations, WWF_Staff_FW, WWF_Staff_WS) %>% 
  mutate(WWF_Staff_FW = ifelse(
    is.na(WWF_Staff_FW),
    0,
    WWF_Staff_FW
  ),
  WWF_Staff_WS = ifelse(
    is.na(WWF_Staff_WS),
    0,
    WWF_Staff_WS
  ))
```

```{r}
rm(df, country_org)
```

## 1.4) Export files

```{r}
rio::export(ngo_country, here("output", "water_orgs", "water_orgs_country.csv"))

write_rds(ngo_country, here("output", "water_orgs", "water_orgs_country.rds"))
```

```{r}
ngo_country %>%
  left_join(country, by = c("country" = "NAME")) %>%
  st_as_sf %>%
  dplyr::select(-c(OBJECTID, ISO_2DIGIT, ISO_3DIGIT)) %>%
  st_write(here("output", "water_orgs", "shp", "water_orgs_country.shp"))
```

# 2) TNC Water Funds

## 2.1) Load and clean data

*as downloaded from https://tnc.app.box.com/s/b5wpi99286t1l1vy8fd7dn62lvj1ur0k*
```{r}
tnc_wf <- read_xlsx(here("input", "water_orgs", "Copy of Tracking all WFs_E.Simmons_list_as of 12-07-2021.xlsx"), 
                    sheet = "ES tracking list of all WFs") %>%
  rename("Region" = 1, 
         "Country" = 2, 
         "CitySite" = 3, 
         "WF_Name" = 4, 
         "Lead" = 5, 
         "Phase" = 6, 
         "Date" = 7, 
         "Status" = 8, 
         "Launch" = 9, 
         "Lat" = 10, 
         "Long" = 11, 
         "In_Map1" = 12, 
         "In_Map2" = 13, 
         "In_Map3" = 14, 
         "In_Map4" = 15, 
         "Links" = 16, 
         "Notes" = 17, 
         "In_Hub" = 18) #%>%
# filter(Status %in%  c("WF Under Development / to be Created", "Water Fund Created", "Water Fund created"))
```

## 2.2) Export files

```{r}
tnc_wf %>%
  filter(!is.na(Lat) & !is.na(Long)) %>%
  st_as_sf(coords = c("Long", "Lat"),
           crs = 4326,
           remove = F) %>%
  st_write(here("output", "water_orgs", "shp", "tnc_wf.shp"), layer_options = "ENCODING=UTF-8", delete_layer = TRUE)
```

# 3) AWS sites

## 3.1) Load and clean data

*as downloaded from https://a4ws.org/certification/certified-sites/ and https://a4ws.org/certification/registered-sites/*
```{r}
aws_cert <- read_csv(here("input", "water_orgs", "List of AWS certified sites - Alliance for Water Stewardship.csv"))

aws_reg <- read_csv(here("input", "water_orgs", "List of AWS registered sites - Alliance for Water Stewardship.csv"))
```

```{r}
aws_cert <- aws_cert %>%
  rename("Regist_No" = 1, 
         "Site" = 2, 
         "Address" = 3, 
         "Country" = 4, 
         "Catchment" = 5, 
         "Cert_Rep" = 6, 
         "Level" = 7,
         "Date" = 8,
         "Valid" = 9,
         "Standard" = 10,
         "CAB" = 11,
         "Industry" = 12,
         "Status" = 13) %>% 
  mutate(AWS = "Certified", .before = 1)
```

```{r}
aws_reg <- aws_reg %>%
  rename("Account" = 1, 
         "Site" = 2, 
         "Address" = 3, 
         "Country" = 4, 
         "Industry" = 5, 
         "Regist_No" = 6, 
         "Legacy_No" = 7) %>% 
  mutate(AWS = "Registered", .before = 1)
```

## 3.2) Geocode

```{r}
aws_sites <- aws_cert %>%
  bind_rows(aws_reg) %>%
  geocode(Address, method = "osm", lat = "Lat" , long = "Long") %>%
  mutate(
    Lat = if_else(Address == "Abu Minqar lat 26.472883 long 27.7680806, Abu Minqar, 0", 26.472883, Lat),
    Long = if_else(Address == "Abu Minqar lat 26.472883 long 27.7680806, Abu Minqar, 0", 27.7680806, Long)
  )
```

## 3.3) Export files

```{r}
aws_sites %>%
  filter(!is.na(Lat) & !is.na(Long)) %>%
  st_as_sf(coords = c("Long", "Lat"),
           crs = 4326,
           remove = F) %>%
  st_write(here("output", "water_orgs", "shp", "aws_sites.shp"), layer_options = "ENCODING=UTF-8", delete_layer = TRUE)
```

# 4) AWS Impact Accelerator

## 4.1) Create table

```{r}
aws_ia <- tibble(country = c("Bangladesh", 
                           "China", 
                           "China", 
                           "China", 
                           "China", 
                           "China",
                           "China",
                           "Chile",
                           "Chile",
                           "Chile",
                           "India",
                           "India",
                           "India",
                           "India",
                           "India",
                           "Pakistan",
                           "Pakistan",
                           "Peru",
                           "Peru",
                           "Spain",
                           "Spain",
                           "Turkey",
                           "Vietnam",
                           "Vietnam"),
               city = c("Dhaka", 
                        "Chengdu", 
                        "Jinan", 
                        "Shanghai", 
                        "Suzhou", 
                        "Wuxi",
                        "Wuhan",
                        "Santiago",
                        "San Felipe",
                        "Quilpué",
                        "Bangalore", 
                        "Chennai", 
                        "Coimbatore", 
                        "Hyderabad", 
                        "Tiruppur",
                        "Lahore",
                        "Karachi",
                        "Olmos",
                        "Ica",
                        "Huelva",
                        "Murcia",
                        "Istanbul",
                        "Hanoi",
                        "Ho Chi Minh City"
                        )) 
```

## 4.2) Geocode

```{r}
aws_ia <- geo(city = aws_ia$city, country = aws_ia$country, method = "osm", full_results = TRUE) %>% 
  select(city, country, lat, long)
```

## 4.3) Export files

```{r}
aws_ia %>%
  drop_na() %>%
  st_as_sf(crs = 4326,
           coords = c("long", "lat"),
           remove = F) %>%
  st_write(here("output", "water_orgs", "shp", "aws_ia.shp"), layer_options = "ENCODING=UTF-8", delete_layer = TRUE)
```

# 5) WaterAid

## 5.1) Load and clean data

```{r}
water_aid <- read_xlsx(here("input", "water_orgs", "District Breakdown of WaterAid Operations.xlsx"))  %>% 
  clean_names() %>% 
  dplyr::select(-c(wa_federation_members,x2, country_programmes)) %>% 
  rename_with(~str_remove(.x, "_limited_work"), contains("_limited_work")) %>% 
  rename(eswatini = e_swatini) %>% 
  pivot_longer(everything(), names_to = "country", values_to = "geo") %>% 
  separate(geo, into = c("geo1", "geo2", "geo3"), sep = ",") %>% 
  pivot_longer(cols = geo1:geo3) %>% 
  dplyr::select(-name) %>% 
  drop_na %>% 
  rename(geo = value) %>%
  mutate(geo = geo %>% str_squish,
         country = country %>% str_to_title) %>% 
  distinct()
```

## 5.2) Geocode

```{r}
temp <- geo(city = water_aid$geo, country = water_aid$country, method = "osm", full_results = TRUE) %>% 
  select(city, country, lat, long)

temp_1 <- temp %>% 
  drop_na

temp_na_city <- temp %>% 
  filter(is.na(lat)) %>% 
  select(city, country)

temp_county <- geo(county = temp_na_city$city, country = temp_na_city$country, method = "osm", full_results = TRUE)

temp_2 <- temp_county %>% 
  drop_na %>% 
  select(county, country)

temp_na_county <- temp_county %>% 
  filter(is.na(lat)) %>% 
  select(county, country)

temp_state <- geo(state = temp_na_county$county, country = temp_na_county$country, method = "osm", full_results = TRUE)  

temp_3 <- temp_state %>% 
  drop_na %>% 
  select(state, country)

places_coord <- read_xlsx(here("input", "water_orgs", "places_coord.xlsx")) %>% # coordinates based on chatGPT
  separate("place", c("place", "country"), sep = ",")

water_aid <- temp_1 %>% 
  rename(place = city) %>% 
  bind_rows(temp_2 %>% rename(place = county)) %>% 
  bind_rows(temp_3 %>% rename(place = state)) %>% 
  bind_rows(places_coord)
```

## 5.3) Export files

```{r}
water_aid %>%
  drop_na() %>%
  st_as_sf(crs = 4326,
           coords = c("long", "lat"),
           remove = F) %>%
  st_write(here("output", "water_orgs", "shp", "water_aid.shp"), layer_options = "ENCODING=UTF-8", delete_layer = TRUE)
```

# 6) wrap

This was compiled manually based on info from wrap. The shp can be found here("output", "water_orgs", "shp", "wrap.shp")
