---
title: "Geographies summarise"
toc: true
format:
  html: 
    code-fold: true
    theme: simplex
---

# 0) Load packages and data

**Load packages**

```{r, warning=FALSE, message=FALSE}
pacman::p_load(tidyverse, here, janitor, sf, skimr, BAMMtools)

sf::sf_use_s2(FALSE) # This is needed because the HydroBASINS data has some geometries issues
```

**Load data**

```{r, warning=FALSE, message=FALSE, results='hide'}
assets <- read_rds(here("input", "assets_cleaned.rds")) %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326, remove = F)

geographies <- list.files(path = here("input", "geographies"), full.names = T, pattern = ".shp$") %>% 
  map(st_read)

names(geographies) <- c("hybas6", "wmo", "country", "adm1")

regions <- read_csv(here("input", "geographies", "world-regions-according-to-the-world-bank.csv")) %>% 
  clean_names()

country_regions <- geographies$country %>% 
  left_join(regions, by = c("ISO_3DIGIT" = "code")) %>% 
  dplyr::select(-c(entity, year)) %>% 
  rename(region = world_region_according_to_the_world_bank) %>% 
  relocate(region, .after = NAME) %>%
  mutate(region = case_when(
    NAME == "Anguilla" ~ "Europe and Central Asia",
    NAME == "Saint Barthelemy" ~ "Europe and Central Asia",
    NAME == "Bonaire" ~ "Europe and Central Asia",
    NAME == "Saba" ~ "Europe and Central Asia",
    NAME == "Saint Eustatius" ~ "Europe and Central Asia",
    #NAME == "Bouvet Island" ~ 
    NAME == "Cocos Islands" ~ "East Asia and Pacific", 
    NAME == "Cook Islands" ~ "East Asia and Pacific",                           
    #NAME == "Clipperton" ~                               
    NAME == "Christmas Island" ~ "East Asia and Pacific",                          
    NAME == "Aksai Chin (disputed)" ~ "East Asia and Pacific",                     
    NAME == "Abyei (disputed)" ~ "Sub-Saharan Africa",                        
    NAME == "Western Sahara (disputed)" ~ "Middle East and North Africa",                
    NAME == "Arunachal Pradesh (disputed)" ~ "South Asia",             
    NAME == "Canarias" ~ "Europe and Central Asia",                                
    NAME == "Falkland Islands" ~ "Latin America and Caribbean",                          
    NAME == "Micronesia" ~ "East Asia and Pacific",                               
    NAME == "French Guiana" ~ "Europe and Central Asia",                           
    NAME == "Guernsey" ~ "Europe and Central Asia",                                 
    NAME == "Guadeloupe" ~ "Europe and Central Asia",                               
    #NAME == "South Georgia and South Sandwich Islands" ~ 
    #NAME == "Heard Island and McDonald Islands" ~         
    #NAME == "British Indian Ocean Territory" ~            
    NAME == "Jersey" ~ "Europe and Central Asia",                                  
    NAME == "Martinique" ~ "Europe and Central Asia",                               
    NAME == "Montserrat" ~ "Europe and Central Asia",                                
    NAME == "Norfolk Island" ~ "East Asia and Pacific",                          
    NAME == "Saint Pierre and Miquelon" ~ "North America",                
    NAME == "Pitcairn" ~ "East Asia and Pacific",                                  
    NAME == "Azores" ~ "Europe and Central Asia",                                 
    NAME == "Madeira" ~ "Europe and Central Asia",                                  
    NAME == "Réunion" ~ "Europe and Central Asia",                                  
    #NAME == "Saint Helena" ~                             
    #NAME == "French Southern Territories" ~               
    NAME == "Tokelau" ~ "East Asia and Pacific",                                  
    NAME == "United States Minor Outlying Islands" ~ "East Asia and Pacific",    
    NAME == "Vatican City" ~ "Europe and Central Asia",                             
    NAME == "Wallis and Futuna" ~ "East Asia and Pacific",                        
    NAME == "Mayotte" ~ "Europe and Central Asia",
    TRUE ~ region
  ))
```

# 1) Prepare data

## 1.1) Exclude sites in the oceans

```{r}
assets <- assets %>% 
  drop_na(country) 
```

## 1.2) Shorten industry names

```{r, warning=FALSE, message=FALSE}
assets$industry %>% 
  unique
```

```{r, warning=FALSE, message=FALSE}
assets <- assets %>% 
  mutate(industry = case_when(
    industry == "Agriculture (animal products)" ~ "agri_a",
    industry == "Agriculture (plant products)" ~ "agri_p",
    industry == "Construction Materials" ~ "constru",
    industry == "Electric Energy Production - Geothermal or Combustion (Biomass, Coal, Gas, Nuclear, Oil)" ~ "geotherm",
    industry == "Electric Energy Production - Hydropower" ~ "hydrop",
    industry == "Electric Energy Production - Solar, Wind" ~ "sol_wind",
    industry == "Fishing and aquaculture" ~ "fishing",
    industry == "Health Care, Pharmaceuticals and Biotechnology" ~ "pharma",
    industry == "Hospitality Services" ~ "hospital",
    industry == "Metals & Mining" ~ "mining",
    industry == "Oil, Gas & Consumable Fuels" ~ "oil_gas",
    industry == "Textiles, Apparel & Luxury Good Production" ~ "textile",
    industry == "Transportation Services" ~ "transpo",
    industry == "Water utilities / Water Service Providers" ~ "water",
    industry == "Other" ~ "other"
  ))
```

# 2) Summarize data by different geographies

This includes:

1.  Number of sites by industry

2.  Region-wide natural breaks (Jenks) by industry

3.  Total number of sites

4.  Density of sites

## 2.1) Create a function to summarise data

```{r, warning=FALSE, message=FALSE}
summarise_geo <- function(geography){

# Prepare geography info
id_nam <- geography %>% 
  names %>%
  magrittr::extract(1)

geography <- geography %>% 
  rename(id_geo = 1) %>% 
  rowid_to_column() %>% 
  mutate(id_geo = factor(id_geo)) %>%
  st_join(dplyr::select(country_regions, region), largest = TRUE)

# Join assets to geography, excluding duplicated assets
as_geo <- geography %>% 
  st_join(assets) %>% 
  st_drop_geometry()

as_geo_no_dup <- as_geo %>% 
  dplyr::filter(duplicated == "no")

as_geo_dup <- as_geo %>% 
  dplyr::filter(duplicated == "yes") %>% 
  distinct(id_geo)

as_geo_uniq <- as_geo_no_dup %>% 
  bind_rows(as_geo_dup)

# 1) Number of sites by industry
count_by_ind <- as_geo_uniq %>% 
  drop_na(id) %>% 
  group_by(id_geo) %>% 
  count(industry) %>% 
  ungroup() %>% 
  pivot_wider(names_from = "industry", names_prefix = "n_", values_from =  "n") %>% 
  inner_join(dplyr::select(geography, rowid, id_geo, region) %>% st_drop_geometry)

# 2) Region-wide natural breaks (Jenks) by industry
regions_list <- count_by_ind %>% 
  drop_na(region) %>% 
  split(~region)

breaks_by_ind <- map(1:length(regions_list), function(i) {
  
  region_i <- regions_list[[i]]
  
  # Filter to industries with >=5 assets (because we have 5 breaks)
  ind_5plus_assets <- region_i %>% 
    dplyr::select(-c(rowid, id_geo, region)) %>% 
    summarise(across(everything(), ~sum(!is.na(.)))) %>% 
    pivot_longer(everything()) %>% 
    dplyr::filter(value >= 5) %>% 
    pull(name)
  
  # Filter to industries with >=4 different values (this is needed to obtain the breaks)
  ind_4plus_diff_values <- region_i %>% 
    dplyr::select(all_of(ind_5plus_assets)) %>% 
    summarise(across(
      everything(), ~n_distinct(.))
    ) %>% 
    pivot_longer(everything()) %>% 
    dplyr::filter(value >= 4) %>% 
    pull(name)
  
  # Get breaks
  breaks <- region_i %>% 
    dplyr::select(all_of(ind_4plus_diff_values)) %>% 
    summarise(across(all_of(ind_4plus_diff_values), ~getJenksBreaks(., k = 5))) %>% 
    rename_with(~str_remove(., "n_"))
  
  # Handle industries with no breaks
  all_ind <- unique(assets$industry)
  
  missing_ind <- setdiff(all_ind, str_remove(ind_4plus_diff_values, "n_")) 
    
  missing_ind <- as_tibble(matrix(NA, ncol = length(missing_ind), nrow = 1)) %>%
    set_names(missing_ind)
  
  # Complete table of breaks
  breaks <- breaks %>% 
    bind_cols(missing_ind)
  
  # Assign breaks
  region_i %>% 
    mutate(b_constru = 
             case_when(
               between(n_constru, breaks$constru[4], breaks$constru[5]) ~ 4,
               between(n_constru, breaks$constru[3], breaks$constru[4]) ~ 3,
               between(n_constru, breaks$constru[2], breaks$constru[3]) ~ 2,
               between(n_constru, breaks$constru[1], breaks$constru[2]) ~ 1),
           b_hydrop = 
             case_when(
               between(n_hydrop, breaks$hydrop[4], breaks$hydrop[5]) ~ 4,
               between(n_hydrop, breaks$hydrop[3], breaks$hydrop[4]) ~ 3,
               between(n_hydrop, breaks$hydrop[2], breaks$hydrop[3]) ~ 2,
               between(n_hydrop, breaks$hydrop[1], breaks$hydrop[2]) ~ 1),
           b_geotherm = 
             case_when(
               between(n_geotherm, breaks$geotherm[4], breaks$geotherm[5]) ~ 4,
               between(n_geotherm, breaks$geotherm[3], breaks$geotherm[4]) ~ 3,
               between(n_geotherm, breaks$geotherm[2], breaks$geotherm[3]) ~ 2,
               between(n_geotherm, breaks$geotherm[1], breaks$geotherm[2]) ~ 1),
           b_sol_wind = 
             case_when(
               between(n_sol_wind, breaks$sol_wind[4], breaks$sol_wind[5]) ~ 4,
               between(n_sol_wind, breaks$sol_wind[3], breaks$sol_wind[4]) ~ 3,
               between(n_sol_wind, breaks$sol_wind[2], breaks$sol_wind[3]) ~ 2,
               between(n_sol_wind, breaks$sol_wind[1], breaks$sol_wind[2]) ~ 1),
           b_mining = 
             case_when(
               between(n_mining, breaks$mining[4], breaks$mining[5]) ~ 4,
               between(n_mining, breaks$mining[3], breaks$mining[4]) ~ 3,
               between(n_mining, breaks$mining[2], breaks$mining[3]) ~ 2,
               between(n_mining, breaks$mining[1], breaks$mining[2]) ~ 1),
           b_oil_gas = 
             case_when(
               between(n_oil_gas, breaks$oil_gas[4], breaks$oil_gas[5]) ~ 4,
               between(n_oil_gas, breaks$oil_gas[3], breaks$oil_gas[4]) ~ 3,
               between(n_oil_gas, breaks$oil_gas[2], breaks$oil_gas[3]) ~ 2,
               between(n_oil_gas, breaks$oil_gas[1], breaks$oil_gas[2]) ~ 1),
           b_textile = 
             case_when(
               between(n_textile, breaks$textile[4], breaks$textile[5]) ~ 4,
               between(n_textile, breaks$textile[3], breaks$textile[4]) ~ 3,
               between(n_textile, breaks$textile[2], breaks$textile[3]) ~ 2,
               between(n_textile, breaks$textile[1], breaks$textile[2]) ~ 1),
           b_transpo = 
             case_when(
               between(n_transpo, breaks$transpo[4], breaks$transpo[5]) ~ 4,
               between(n_transpo, breaks$transpo[3], breaks$transpo[4]) ~ 3,
               between(n_transpo, breaks$transpo[2], breaks$transpo[3]) ~ 2,
               between(n_transpo, breaks$transpo[1], breaks$transpo[2]) ~ 1),
           b_pharma = 
             case_when(
               between(n_pharma, breaks$pharma[4], breaks$pharma[5]) ~ 4,
               between(n_pharma, breaks$pharma[3], breaks$pharma[4]) ~ 3,
               between(n_pharma, breaks$pharma[2], breaks$pharma[3]) ~ 2,
               between(n_pharma, breaks$pharma[1], breaks$pharma[2]) ~ 1),
           b_other = 
             case_when(
               between(n_other, breaks$other[4], breaks$other[5]) ~ 4,
               between(n_other, breaks$other[3], breaks$other[4]) ~ 3,
               between(n_other, breaks$other[2], breaks$other[3]) ~ 2,
               between(n_other, breaks$other[1], breaks$other[2]) ~ 1),
           b_agri_a = 
             case_when(
               between(n_agri_a, breaks$agri_a[4], breaks$agri_a[5]) ~ 4,
               between(n_agri_a, breaks$agri_a[3], breaks$agri_a[4]) ~ 3,
               between(n_agri_a, breaks$agri_a[2], breaks$agri_a[3]) ~ 2,
               between(n_agri_a, breaks$agri_a[1], breaks$agri_a[2]) ~ 1),
           b_agri_p = 
             case_when(
               between(n_agri_p, breaks$agri_p[4], breaks$agri_p[5]) ~ 4,
               between(n_agri_p, breaks$agri_p[3], breaks$agri_p[4]) ~ 3,
               between(n_agri_p, breaks$agri_p[2], breaks$agri_p[3]) ~ 2,
               between(n_agri_p, breaks$agri_p[1], breaks$agri_p[2]) ~ 1),
           b_fishing = 
             case_when(
               between(n_fishing, breaks$fishing[4], breaks$fishing[5]) ~ 4,
               between(n_fishing, breaks$fishing[3], breaks$fishing[4]) ~ 3,
               between(n_fishing, breaks$fishing[2], breaks$fishing[3]) ~ 2,
               between(n_fishing, breaks$fishing[1], breaks$fishing[2]) ~ 1),
           b_water = 
             case_when(
               between(n_water, breaks$water[4], breaks$water[5]) ~ 4,
               between(n_water, breaks$water[3], breaks$water[4]) ~ 3,
               between(n_water, breaks$water[2], breaks$water[3]) ~ 2,
               between(n_water, breaks$water[1], breaks$water[2]) ~ 1),
           b_hospital = 
             case_when(
               between(n_hospital, breaks$hospital[4], breaks$hospital[5]) ~ 4,
               between(n_hospital, breaks$hospital[3], breaks$hospital[4]) ~ 3,
               between(n_hospital, breaks$hospital[2], breaks$hospital[3]) ~ 2,
               between(n_hospital, breaks$hospital[1], breaks$hospital[2]) ~ 1),
    )
}) %>% 
  # Put lists (individual regions) back together
  bind_rows() %>%

# 3) Total number of sites
  rowwise() %>%
  mutate(n_sites = sum(c_across(dplyr::select(count_by_ind, -c(id_geo, region, rowid)) %>% names), na.rm = T)) %>% 
  ungroup()

# Join pieces together  
geo_summary <- geography %>% 
  left_join(breaks_by_ind, by = "id_geo") %>%
  dplyr::select(
    id_geo,
    area_km2 = Area_km2,
    region = region.x,
    starts_with("n_"),
    starts_with("b_")
  ) %>%
  mutate(d_sites = n_sites/area_km2, .after = n_sites) %>% # 4) Density of sites
  relocate(n_sites:d_sites, .after = region) %>%
  rename_with(~get("id_nam"), id_geo)
}
```

## 2.2) Apply function to the geography

```{r, warning=FALSE, message=FALSE}
geo_summary <- summarise_geo(geographies$hybas6)
```

```{r}
geo_summary %>% 
  st_drop_geometry() %>%
  skim
```

# 3) Export files

```{r}
write_rds(geo_summary, here("output", "summarise_geo", "1_summarise", "hybas6_summary.rds"))
```
