---
title: "Geographies enrichment"
toc: true
format:
  html: 
    code-fold: true
    theme: simplex
---

# 0) Load packages and data

```{r}
pacman::p_load(tidyverse, here, janitor, sf, skimr, exactextractr, raster, terra)
```

**Load data**

```{r, warning=FALSE, message=FALSE, results='hide'}
gdp <- raster(here("input", "raster_gdp_pop", "GDP_PPP_2015.tif"))

pop <- raster(here("input", "raster_gdp_pop", "ppp_2020_1km_Aggregated.tif"))

ci <- rast(here("input", "raster_GCI30", "GCI30_agg.tif"))

vp <- rast(here("input", "raster_value_crop_prod", "spam2010V2r0_global_V_agg_VP_CROP_A.tif"))

rfs <- st_read(here("input", "rfs_data", "rfs_data.shp"))

country_adm1 <- st_read(here("input", "geographies", "world_adm1_esri_edited_diss.shp"))

wmo <- st_read(here("input", "geographies", "wmobb_basins.shp"))
```

# 1) Extract GDP and population values

```{r}
ext_gdp_pop <- function(geo){
  
  gdp_geo <- exact_extract(gdp, geo, fun = "sum")
  
  pop_geo <- exact_extract(pop, geo, fun = "sum")

  geo %>%
    mutate(
      gdp_sum = gdp_geo/1000000,
      gdp_dens = gdp_sum/area_km2,
      pop_sum = pop_geo,
      pop_dens = pop_sum/area_km2
  ) %>%
    dplyr::select(1, gdp_sum:pop_dens) %>%
    st_drop_geometry()
}
```

# 2) Extract land cover (MODIS) percentage by geography
**Temporarily remove to speed things up**
# 3) Extract wetland type (CIFOR) percentage by geography
**Temporarily remove to speed things up**
# 4) Extract agri commodity (MapSpam) area sum by Geography
**Temporarily remove to speed things up**
# 5) Extract cropping intensity (GCI30)

```{r}
ext_ci <- function(geo){
  
  ci_geo <- exact_extract(ci, geo, fun = "mean")

  geo %>% 
    dplyr::select(1) %>%
    mutate(ci_mean = ci_geo) %>%
    st_drop_geometry()
}
```

# 6) Extract value of production (MapSpam)

```{r}
ext_vp <- function(geo){
  
  vp_geo <- exact_extract(vp, geo, fun = "mean")

  geo %>% 
    dplyr::select(1) %>%
    mutate(vp_mean = vp_geo) %>%
    st_drop_geometry()
}
```

# 7) Apply functions to geographies

```{r}
enrich_geo <- function(geo_summary){
  
  geo <- read_rds(here("output", "summarise_geo", "1_summarise", geo_summary))
  
  geo %>%
    left_join(ext_gdp_pop(geo)) %>%
    left_join(ext_ci(geo)) %>%
    left_join(ext_vp(geo)) %>%
    mutate(across(where(is.double), ~ round(., 2))) %>%
    st_join(dplyr::select(country_adm1, country = COUNTRY, adm1 = NAME), largest = TRUE) %>%
    st_join(dplyr::select(wmo, wmobb_id = WMOBB, wmo_basin = WMOBBName), largest = TRUE) %>%
    relocate(country:wmo_basin, .after = region) %>%
    st_join(dplyr::select(rfs, RC1, RC2, RC3, S2_4, RC8, Phy_P3rc), largest = TRUE)
}
```

```{r}
# hybas7_enriched <- enrich_geo("hybas6_summary.rds")
# 
# hybas6_enriched <- enrich_geo("hybas6_summary.rds")

hybas5_enriched <- enrich_geo("hybas5_summary.rds")
```

```{r}
hybas5_enriched %>% 
  st_drop_geometry() %>%
  skim
```

# 8) Export files

```{r}
# write_rds(hybas7_enriched, here("output", "summarise_geo", "2_enrichment", "final", "hybas7_enriched.rds"))
# 
# write_rds(hybas6_enriched, here("output", "summarise_geo", "2_enrichment", "final", "hybas6_enriched.rds"))

write_rds(hybas5_enriched, here("output", "summarise_geo", "2_enrichment", "final", "hybas5_enriched.rds"))
```
